Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-12 arxiv/experiments.tex
revise
fix grammar

\paragraph{Sample efficiency.}
First, we study the setting of Example \ref{ex_same_cov}.
Figure \ref{fig_same_cov} shows the average prediction loss as we increase the number of samples per-task   from $400$ to $2000$.
In all the parameter settings, our results estimate the empirical losses accurately.
We also observe a trend where the average prediction loss increases as we increase $d$ from $0.1$ to $0.2$.
Our work explains the differences between these two settings since $d^2 = 0.1^2$ is always smaller than $\frac{\sigma^2 p}{n - p}$, but $d^2 = 0.2^2$ is not.
%Second, when $d = 0.1$, we have that $d^2 \le \frac{\sigma^2 p}{n - p}$ for all values of $n$, hence the average prediction loss of hard parameter sharing is always lower than STL.
Indeed, we observe a crossover point between hard parameter sharing and STL.
Finally, looking horizontally at the same prediction loss level, we find that hard parameter sharing requires 3x fewer samples per-task than STL. \FY{I do not quite understand this sentence about "3x fewer", because how much data needed depends on $d$ and the prediction loss level we are looking at. For example, at the cross point, this ratio is 1. }

>>>>

\paragraph{Sample efficiency.}
First, we study the setting of Example \ref{ex_same_cov}.
Figure \ref{fig_same_cov} shows the average prediction loss as we increase the number of samples per-task   from $400$ to $2000$.
In all the parameter settings, our results estimate the empirical losses accurately.
We also observe a trend where the average prediction loss increases as we increase $d$ from $0.1$ to $0.2$.
Our work explains the differences between these two settings since $d^2 = 0.1^2$ is always smaller than $\frac{\sigma^2 p}{n - p}$, but $d^2 = 0.2^2$ is not.
%Second, when $d = 0.1$, we have that $d^2 \le \frac{\sigma^2 p}{n - p}$ for all values of $n$, hence the average prediction loss of hard parameter sharing is always lower than STL.
Indeed, we observe a crossover point between hard parameter sharing and STL.
Finally, for $d = 0.2$, looking horizontally at the same prediction loss level, we find that hard parameter sharing requires fewer samples per-task than STL. %\FY{I do not quite understand this sentence about "3x fewer", because how much data needed depends on $d$ and the prediction loss level we are looking at. For example, at the cross point, this ratio is 1. }

### 2020-10-12 arxiv/setup.tex
fix grammar

Regression analysis is arguably one of the most fundamental methods in statistics and machine learning.
In many applications of multi-task learning, the amount of labeled data from each dataset is usually insufficient for learning a single task.
For example, this is the case if a dataset's sample size is only a small constant times the dimension of the feature vectors \cite{GLUE}.
In this paper, we consider multi-task learning in a high-dimensional linear regression setting.
We assume that there are multiple datasets that all follow a (potentially different) linear model.
In each dataset, we assume the feature vector $x = {\Sigma}^{1/2} z$, where $z \in \real^p$ has i.i.d entries with mean zero, unit variance, and covariance $\Sigma \in\real^{p \times p}$ that is deterministic and positive semidefinite.
We are mainly interested in the \textit{high-dimensional} setting, where each dataset's sample size and the feature dimension grow linearly at a fixed ratio.
This setting is particularly challenging, but is crucial for understanding how datasets' properties such as their sample sizes impact generalization performance.

We study a hard parameter sharing architecture to learn jointly from multiple datasets.
In this architecture, there is a shared feature representation layer $B\in\real^{p\times r}$ for all tasks, and a task-specific output layer $A_i \in \real^r$ for every task $i$ to $t$.
Suppose we have $t$ datasets.
For each task $i = 1, 2, \dots, t$, let $n_i$ denote its sample size, $X^{(i)} \in \real^{n_i \times p}$ denote its feature covariates, and $Y^{(i)} \in \real^{n_i}$ denote all the labels.
%The width of $B$, denoted by $r$, plays an important role in regularization.
%As observed in Proposition 1 of \citet{WZR20}, if $r \ge t$, there is no regularization effect.
%Hence, we assume that $r < t$ in our study.
%For example, when there are only two tasks, $r = 1$ and $B$ reduces to a vector whereas $W_1, W_2$ become scalars.
Let $A = [A_1, A_2, \dots, A_t] \in \real^{r \times t}$.
We study the following optimization objective.
%	\item Separate each dataset $(X_i, Y_i)$ randomly into a training set $(X_i^{tr}, Y_i^{tr})$ and a validation set $(X_i^{val}, Y_i^{val})$.
%	The size of each set is described below.
%	\item Learn the shared layer $B$: minimize the training loss over $B$ and $W_1, \dots, W_t$, leading to a local minimum of $B$ that depends on $W_1, \dots, W_t$, denoted by $\hat{B} = \hat{B}(W_1, \dots, W_t)$.
\begin{align}\label{eq_mtl}
			f(A, B) = \sum_{i=1}^t \norm{X^{(i)} B A_i - Y^{(i)}}^2.
\end{align}
Given a solution $\hat{A}, \hat{B}$ from the above optimization, let $\hat{\beta}_i^{\MTL} = \hat{B} \hat{A}_i$ denote the hard parameter sharing (HPS) estimator for task $i$.
The important questions here are:
(i) how well does the estimator work? In particular, how does the performance of the estimator scale with sample size?
(ii) for datasets with different sample sizes and covariate shifts, how do they impact the estimator?

@@

\paragraph{Main results.}
Our first result (Theorem \ref{thm_many_tasks}) applies to the multi-label prediction setting where all datasets have the same features, and we want to make multiple predictions (e.g. \cite{hsu2009multi}).
We analyze the global minimum of $f(A, B)$, and provide a sharp generalization bound for its (out-of-sample) prediction loss for any task.
This is a tractable special case of the general problem because in the worst case, $f(A, B)$ is non-convex in $A$ and $B$ (e.g. matrix completion is a special case for suitably designed $X^{(i)}, Y^{(i)}$).
We show that the prediction loss of hard parameter sharing admits a clean bias-variance decomposition.
Our results imply that hard parameter sharing helps by reducing variance compared to single-task learning, but hurts by increasing bias.
Our second result (Theorem \ref{thm_main_RMT}) applies to two tasks with arbitrarily different sample sizes and covariance matrices.
We analyze the local minimum of $f(A, B)$, and provide a sharp generalization bound for its prediction loss for both tasks.
Despite its simplicity, we show several rich phenomena by varying sample sizes and covariate shifts in this setting.
See Figure \ref{fig_intro_sample_size} for an illustration.

\begin{figure}[!t]
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.65\textwidth]{figures/mtl_model.pdf}
		\caption{A hard parameter sharing architecture}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.9\textwidth]{figures/sample_ratio_c2_400.eps}
		\caption{Varying sample size ratio}
		\label{fig_intro_sample_size}
	\end{subfigure}
	\caption{An illustrative example of our result:
	Consider the prediction loss of hard parameter sharing (left) for task two, given two linear regression tasks.
	Increasing task one's sample size helps reduce task two's prediction loss initially, but increases afterwards. Our result provides an estimated loss (solid line) that matches the empirical loss (dotted line) accurately. This phenomenon occurs due to different bias-variance tradeoffs as the sample size ratio increases.
	See Section \ref{sec_simulation} for the precise setting.}
	\label{fig_intro_sample_size}
\end{figure}

As a consequence of our concrete and precise loss estimates, we observe qualitative properties of hard parameter sharing for varying datasets' properties.
\begin{enumerate}
	\item \textit{Sample efficiency (Example \ref{ex_same_cov})}:
	One advantage of combining multiple datasets is that the requirement for labeled data reduces compared to STL, a phenomenon that is empirically observed in Taskonomy \cite{ZSSGM18}.
	Our results further show that HPS's sample efficiency depends on task-specific variance vs. noise variance. It is generally high when the noise variance is large compared to model-specific variance across tasks.
	\item \textit{Sample size ratio (Example \ref{ex_sample_ratio})}: Increasing one task's sample size does not always help reduce another task's loss. We find in a simplified setting that depending on how much bias increases, the task loss either decreases first before increasing afterwards, or decreases monotinically. These two trends result from different tradeoffs between increasing bias and decreasing variance.
	\item \textit{Covariate shift (Example \ref{ex_covshift})}: In addition to sample sizes, variance also scales with the covariate shift between different datasets. For high sample size ratio, variance is smallest when there is no covariate shift. Counterintuitively, for low sample size ratio, having covariate shift reduces variance through a complementary spectrum.
\end{enumerate}

@@

There are two main ideas in our analysis. The proof of our first result is based on a geometric intuition that hard parameter sharing finds a ``rank-$r$'' approximation of the datasets.
We carefully keep track of the concentration error between the global minimum of $f(A, B)$ and their population version.
The proof of our second result is significantly more involved because of different sample sizes and covariate shifts.
\HZ{add several sentences on the technical insight}
One limitation of our analysis is that for two tasks with different sample sizes, there is extra error term in the prediction loss of hard parameter sharing (cf. equation \eqref{cor_MTL_error}), which can be large for very small $n_1$.
\HZ{add why this is challenging}

Finally, we discuss the practical implications of work.
One natural implication from our sample size ratio study is a progressive training procedure that gradually adds more data until performance drops.
For example, in the setting of Figure \ref{fig_intro_sample_size}, this procedure will stop right at the minimum of the local basin.
We conduct further studies of this procedure on six text classification datasets, and observe that it reduces the computational cost by $65\%$ compared to a standard round-robin training procedure, while keeping average accuracy of all tasks at the same level.

>>>>

Regression analysis is arguably one of the most fundamental methods in statistics and machine learning.
In many multi-task learning applications, the amount of labeled data from each dataset is usually insufficient for learning a single task.
For example, this is the case if a dataset's sample size is only a small constant factor of the feature dimension \cite{GLUE}.
In this paper, we consider multi-task learning in a high-dimensional linear regression setting.
We assume that there are multiple datasets that all follow a (potentially different) linear model.
In each dataset, suppose the feature vector $x = {\Sigma}^{1/2} z$, where $z \in \real^p$ has i.i.d entries with mean zero, unit variance, and covariance $\Sigma \in\real^{p \times p}$ that is deterministic and positive semidefinite.
We are mainly interested in the \textit{high-dimensional} setting, where each dataset's sample size and the feature dimension grow linearly at a fixed ratio.
This setting is challenging but is crucial for understanding how datasets' sample sizes impact generalization performance.

We study a hard parameter sharing architecture to learn jointly from multiple datasets.
In this architecture, there is a shared feature representation layer $B\in\real^{p\times r}$ for all tasks and a task-specific output layer $A_i \in \real^r$ for every task $i$ to $t$.
Suppose we have $t$ datasets.
For each task $i = 1, 2, \dots, t$, let $n_i$ denote its sample size, $X^{(i)} \in \real^{n_i \times p}$ denote its feature covariates, and $Y^{(i)} \in \real^{n_i}$ denote all the labels.
%The width of $B$, denoted by $r$, plays an important role in regularization.
%As observed in Proposition 1 of \citet{WZR20}, if $r \ge t$, there is no regularization effect.
%Hence, we assume that $r < t$ in our study.
%For example, when there are only two tasks, $r = 1$ and $B$ reduces to a vector whereas $W_1, W_2$ become scalars.
Let $A = [A_1, A_2, \dots, A_t] \in \real^{r \times t}$.
We study the following optimization objective.
%	\item Separate each dataset $(X_i, Y_i)$ randomly into a training set $(X_i^{tr}, Y_i^{tr})$ and a validation set $(X_i^{val}, Y_i^{val})$.
%	The size of each set is described below.
%	\item Learn the shared layer $B$: minimize the training loss over $B$ and $W_1, \dots, W_t$, leading to a local minimum of $B$ that depends on $W_1, \dots, W_t$, denoted by $\hat{B} = \hat{B}(W_1, \dots, W_t)$.
\begin{align}\label{eq_mtl}
			f(A, B) = \sum_{i=1}^t \norm{X^{(i)} B A_i - Y^{(i)}}^2.
\end{align}
Given a solution $\hat{A}, \hat{B}$ from the above optimization, let $\hat{\beta}_i^{\MTL} = \hat{B} \hat{A}_i$ denote the hard parameter sharing (HPS) estimator for task $i$.
The critical questions here are:
(i) how well does the estimator work? In particular, how does the performance of the estimator scale with sample size?
(ii) for datasets with different sample sizes and covariate shifts, how do they impact the estimator?

@@

\paragraph{Main results.}
Our first result (Theorem \ref{thm_many_tasks}) applies to the multi-label prediction setting where all datasets have the same features, and we want to make multiple predictions (e.g. \cite{hsu2009multi}).
We analyze the global minimum of $f(A, B)$, and provide a sharp generalization bound for its (out-of-sample) prediction loss for any task.
This case is tractable even though in general, $f(A, B)$ is non-convex in $A$ and $B$ (e.g. matrix completion is a special case for suitably designed $X^{(i)}, Y^{(i)}$).
We show that the prediction loss of hard parameter sharing admits a clean bias-variance decomposition.
Our results imply that hard parameter sharing helps by reducing variance compared to single-task learning, but hurts by increasing bias.
Our second result (Theorem \ref{thm_main_RMT}) applies to two tasks with arbitrarily different sample sizes and covariance matrices.
We analyze the local minimum of $f(A, B)$ and provide a sharp generalization bound for both tasks' prediction loss.
Despite its simplicity, we show several rich phenomena by varying sample sizes and covariate shifts in this setting.
See Figure \ref{fig_intro_sample_size} for an illustration.

\begin{figure}[!t]
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.65\textwidth]{figures/mtl_model.pdf}
		\caption{A hard parameter sharing architecture}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.9\textwidth]{figures/sample_ratio_c2_400.eps}
		\caption{Varying sample size ratio}
		\label{fig_intro_sample_size_b}
	\end{subfigure}
	\caption{An illustrative example of our result:
	Consider the prediction loss of hard parameter sharing (left) for task two, given two linear regression tasks.
	Increasing task one's sample size helps reduce task two's prediction loss initially, but increases afterward. Our result provides an estimated loss (solid line) that accurately matches the empirical loss (dotted line). This phenomenon occurs due to different bias-variance tradeoffs as the sample size ratio increases.
	See Section \ref{sec_simulation} for the precise setting.}
	\label{fig_intro_sample_size}
\end{figure}

Consequently, we observe qualitative properties of hard parameter sharing for varying datasets' properties using our precise loss estimates.
\begin{enumerate}
	\item \textit{Sample efficiency (Example \ref{ex_same_cov})}:
	One advantage of combining multiple datasets is that the requirement for labeled data reduces compared to STL, a phenomenon that \citet{ZSSGM18} has observed empirically observed.
	Our results further show that HPS's sample efficiency depends on task-specific variance vs. noise variance. It is generally high when the noise variance is large compared to model-specific variance across tasks.
	\item \textit{Sample size ratio (Example \ref{ex_sample_ratio})}: Increasing one task's sample size does not always help reduce another task's loss. In a simplified setting, we find that the task loss either decreases first before increasing afterward or decreases monotonically depending on how much bias increases. These two trends result from different tradeoffs between increasing bias and decreasing variance.
	\item \textit{Covariate shift (Example \ref{ex_covshift})}: In addition to sample sizes, variance also scales with the covariate shift between different datasets. For a high sample size ratio, HPS's  variance is smallest when there is no covariate shift. Counterintuitively, for a low sample size ratio, having covariate shifts reduces variance through a complementary spectrum.
\end{enumerate}

@@

There are two main ideas in our analysis. The proof of our first result uses a geometric intuition that hard parameter sharing finds a ``rank-$r$'' approximation of the datasets.
We carefully keep track of the concentration error between the global minimum of $f(A, B)$ and their population version.
The proof of our second result is significantly more involved because of different sample sizes and covariate shifts.
\HZ{add several sentences on the technical insight}
One limitation of our analysis is that for two tasks with different sample sizes, there is extra error term in the prediction loss of hard parameter sharing (cf. equation \eqref{cor_MTL_error}), which can be large for very small $n_1$.
\HZ{add why this is challenging}

Finally, we discuss the practical implications of work.
Our sample size ratio study implies a concrete progressive training procedure that gradually adds more data until performance drops.
For example, in the setting of Figure \ref{fig_intro_sample_size_b}, this procedure will stop right at the minimum of the local basin.
We conduct further studies of this procedure on six text classification datasets and observe that it reduces the computational cost by $65\%$ compared to a standard round-robin training procedure while keeping the average accuracy of all tasks simultaneously.
